---
title: "Regression diagnostics"
author: "Jerome Hilaire"
output:
  html_document: default
params:
  data: data.frame()
---

#Initialisation and data processing
This hidden part of the code loads the necessary libraries and performs a simple LM
```{r Initialise, echo=FALSE, warning=FALSE, message=FALSE}
# Taken from: http://www.statmethods.net/stats/rdiagnostics.html
# suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
# suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
# suppressWarnings(suppressPackageStartupMessages(library(tidyr)))
suppressWarnings(suppressPackageStartupMessages(library(car)))
suppressWarnings(suppressPackageStartupMessages(library(gvlma)))
suppressWarnings(suppressPackageStartupMessages(library(MASS)))
library(knitr)
library(markdown)
library(rmarkdown)

if(length(params$data) == 0) stop("Please provide a valid dataset.")

# Save data locally
data  <- params$data
# Identify dependent and independent variables
dvar  <- names(data)[grep("E_CC", names(data))]
idvar <- names(data)[which(!names(data) %in% c("iso", "year", "log_E_CC"))] 
# Compute (multi)linear regression
fit <- lm(
  as.formula(paste(dvar, " ~ ", paste(idvar, collapse=" + "))), 
  data=data)
```



#Searching for outliers
**Bonferroni Outlier Test**. This test reports the Bonferroni p-values for Studentized residuals in linear and generalized linear models, based on a t-test for linear models and normal-distribution test for generalized linear models.
```{r Assessing Outliers 1, echo=FALSE}
# Bonferonni p-value for most extreme obs
tryCatch(
  out1 <- outlierTest(fit, labels = paste0(data$iso, "-", data$year)),
  error = function(e) {
    print("Error: could not perform outlier test. Check presence of NA values.")
  })
if (length(grep("out1", ls())) != 0) {
  print(out1)
  cat(paste0("Outliers (based on Bonf. p-value): ", paste0(names(out1$rstudent), collapse=", "), "\n"))
}
```

**QQ-plot**. This plot shows empirical quantiles of a variable, or of studentized residuals from a linear model, against theoretical quantiles of a comparison distribution.
```{r Assessing Outliers 2, echo=FALSE}
out2 <- qqPlot(fit, main="QQ Plot", id.n=20, labels = paste0(data$iso, "-", data$year)) #qq plot for studentized resid 
cat(paste0("Outliers (based on studentized residuals): ", paste0(names(out2), collapse=", "), "\n"))

```

**Cook's D plot**. This plots helps identifying influencial observations. It reports which D values are greater than 4/(n-k-1) 
```{r Checking for influential Observations, echo=FALSE}
cutoff <- 4/((nrow(data)-length(fit$coefficients)-2)) 
plot(fit, which=4, cook.levels=cutoff)

```


#Cheking the distribution of studentized residuals
**The Studentized residuals**. Like standardized residuals, these are normalized to unit variance, but the Studentized version is fitted ignoring the current data point. (They are sometimes called jackknifed residuals).
```{r Normality of Residuals, echo=FALSE}
sresid <- studres(fit) 
hist(sresid, freq=FALSE, 
     main="Distribution of Studentized Residuals")
xfit<-seq(min(sresid),max(sresid),length=40) 
yfit<-dnorm(xfit) 
lines(xfit, yfit, col="red")
```


#Checking linear model assumptions
##Global test of model assumptions
This function performs a global validation of linear models assumptions.
Note: Heteroskedasticity test will probably fail.
```{r Global test of model assumptions, echo=FALSE}
gvmodel <- gvlma(fit, alphalevel = 0.05) 
summary(gvmodel)
```

##Evaluating homoscedasticity
**Non-constant error variance test**. 
```{r Non-constant error variance test, echo=FALSE}
ncvTest(fit)
```

**Spread-level plots**. These plots are used for examining the possible dependence of spread on level, or an extension of these plots to the studentized residuals from linear models.
```{r Spread-level plots, echo=FALSE}
spreadLevelPlot(fit)
```

##Evaluating Collinearity
**Variance inflation factors**. This function calculates variance-inflation and generalized variance-inflation factors for linear and generalized linear models.
```{r Variance inflation factors, echo=FALSE}
vif(fit) 
vars <- paste0(names(sqrt(vif(fit)) > 2)[which(sqrt(vif(fit)) > 2)], collapse=", ")
if (length(vars) > 0) cat(paste0("VIF results - Potential problems with: ", vars, "\n"))
```



##Evaluating Auto-collinearity
**Durbin-Watson Test**. This function computes residual autocorrelations and generalized Durbin-Watson statistics and their bootstrapped p-values.
```{r Test for Autocorrelated Errors, echo=FALSE}
ace <- durbinWatsonTest(fit, max.lag=10)
print(ace)
```
